---
title: "UCF Bachelor's Thesis"
subtitle: "Analysis of Coded Data"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

# Pilot

Importing the Atlas.ti dataset. (NB: I changed the document number for the second holding of Finogenov to make distinguishing easier)

```{r}
library(readxl)
pilot <- read_excel("pilot coding data.xlsx")
```

Merging the observations spread out over multiple rows into one.

```{r}
pilot2 <- pilot %>% 
  select(D, 3, 13:70) %>% 
  group_by(D) %>% 
  summarise_all(list(~ toString(unique(na.omit(.)))))
```

Getting summary statistics.

```{r}
pilot2 %>% 
  summarise_all(list(~n()))

pilot_subset <- pilot2 %>% 
  select(D, Q49:Q53) %>%  
  pivot_longer(names_to = "var", values_to = "val", cols = c(everything(), -D))

pilot_subset %>% 
  group_by(var, val) %>% 
  summarise_all(list(~length(.)))
```

---

# Tidying the Final Dataset

Importing the Excel coded cases dataset:

```{r}
library(readxl)
cases_raw <- read_excel("case codings database.xlsx")
```

Removing the explanation for each level, turning ordinal variables into ordinal factors, and the remaining ones into normal factors:

```{r}
cases <- cases_raw %>% 
  select(gc_dec:l_train) %>% 
  mutate_all(list(~ str_replace(., "\\s=\\s.*", ""))) %>% 
  mutate(
    scrutiny = scrutiny     %>% as.numeric() %>% as.ordered(),
    f_prev = f_prev         %>% as.numeric() %>% as.ordered(),
    f_dispute = f_dispute   %>% as.numeric() %>% as.ordered(),
    uof_deaths = uof_deaths %>% as.numeric() %>% as.ordered(),
    t_who = t_who           %>% as.numeric() %>% as.ordered(),
    t_nature = t_nature     %>% as.numeric() %>% as.ordered(),
    uof_sa = uof_sa         %>% as.numeric() %>% as.ordered(),
    uof_nature = uof_nature %>% as.numeric() %>% as.ordered(),
    o_control = o_control   %>% as.numeric() %>% as.ordered()
  )  %>% 
  mutate_all(list(~ as_factor(.)))

cases <- cbind(select(cases_raw, ID:respondent), cases)
```

# Collapsing Variables and Categories

```{r}
IVcases <- cases %>% select(scrutiny:l_train) #Only want to analyze and reduce the IVs
```

Checking if there are any superfluous levels.

```{r}
IVcases %>% sapply(table)
IVcases %>% sapply(table) %>% sapply(prop.table)
```

There are multiple methods possible to achieve this. Principally, these can be divided into factor analysis and principal component analysis (PCA). Both have variations that are compatible with categorical data. I pick PCA here because, conceptually, it focuses on reducing the number of dimensions (variables) while factor analysis is focused on uncovering latent factors. Within the realm of non-linear PCA, I picked multiple factor analysis (MFA) because it is well suited for grouped data (like in my case)

```{r}
library(FactoMineR)
library(factoextra)


MFAcases <- MFA(IVcases, group = c(6, 14, 5, 11, 3), #Groups based on coding form
                type = rep("n", 5), 
                name.group = c("assessment", "exceptions", "force", "operation", "law"))
```

Using Chi-square and Cramer's V to find relations of dependence between the cateogrical variables

# Using Classification Trees

Installing `rtemis`, a library for advanced machine learning, used here for the additive trees.

```{r eval=FALSE}
install.packages("remotes")
remotes::install_github("egenn/rtemis")
install.packages(c("e1071", "gbm", "glmnet", "pbapply", "plyr", "ranger", "rpart", "data.tree", "DiagrammeR", "missRanger")) #Additional dependencies
```

Test run:

```{r}
library(rtemis)
```


```{r}
cases_test <- cases %>% select(scrutiny:l_train, h_viol)
# cases_test <- cases_raw %>%
#   select(scrutiny:l_train, h_viol) %>%
#   mutate_all(list(~ str_replace(., "\\s=\\s.*", ""))) %>%
#   mutate_all(list(~ dplyr::recode(.,`-99` = "99",
#                                   `-88` = "88",
#                                    `-77` = "77"
#                                   )))
checkData(cases_test)
cases_test2 <- preprocess(cases_test, character2factor = TRUE)
# write_csv(cases_test2, "cases_test.csv")

# cases_test <- cases
# cases_test$violation <- cases$h_viol
# cases_test2 <- cases_test %>% 
#   mutate(o_evac = recode(o_evac, "0" = "1"), o_distinct = recode(o_distinct, "0" = "1")) %>% 
#   mutate(o_evac = replace_na(o_evac, 0), o_distinct = replace_na(o_distinct, 0)) %>% 
#   preprocess(impute = TRUE, character2factor = TRUE, impute.type = "missForest")


checkData(cases_test2)

cases_test.tree <- s.ADDTREE(cases_test, gamma = 0.5, learning.rate = 0.0001, upsample = TRUE)
dplot3.addtree(cases_test.tree)
print(cases_test.tree$mod$addtree.pruned, "Estimate")

#With testing
res <- resample(cases_test, n.resamples = 77, resampler = "kfold")
cases_test.train <- cases_test[res$Fold_1, ]
cases_test.test <- cases_test[-res$Fold_1, ]
cases_test.tree <- s.ADDTREE(x = cases_test.train, x.test = cases_test.test, gamma = 0.5, learning.rate = 0.0001, upsample = TRUE)
```

